---
title: "EDA"
author: "Elizabeth Ho"
output: pdf_document
---

# Load in data
```{r}
# Load in the data
train <- read.csv("data/train_data.csv")
```

## Check the variable types
```{r}
str(train)
```

```{r}
full_model <- lm(Y ~ ., data = train)
summary(full_model)
```
# Visualizing data distributions to understand the characteristics of variables
(Adding in after observing multicollinearity)

# Single variables
```{r}
library(ggplot2)
# Area
ggplot(train, aes(x = Area)) +
    geom_histogram(bins = 25)
# Perimeter
ggplot(train, aes(x = Perimeter)) +
    geom_histogram(bins = 25)
# MajorAxisLength
ggplot(train, aes(x = MajorAxisLength)) +
    geom_histogram(bins = 25)
ggplot(train, aes(x = MinorAxisLength)) +
    geom_histogram(bins = 25)
```

```{r}
ggplot(train, aes(x = AspectRation)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = Eccentricity)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = ConvexArea)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = EquivDiameter)) +
    geom_histogram(bins = 25)
```

```{r}
ggplot(train, aes(x = Extent)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = Solidity)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = roundness)) +
    geom_histogram(bins = 25)

ggplot(train, aes(x = Compactness)) +
    geom_histogram(bins = 25)
```

```{r}
ggplot(train, aes(x = ShapeFactor1)) +
    geom_histogram(bins = 25)
ggplot(train, aes(x = ShapeFactor2)) +
    geom_histogram(bins = 25)
ggplot(train, aes(x = ShapeFactor3)) +
    geom_histogram(bins = 25)
ggplot(train, aes(x = ShapeFactor4)) +
    geom_histogram(bins = 25)
```
# Multiple Variables
Use: barplot, scatterplot.

## Area vs. perimeter scatterplot
```{r, warning=FALSE}
library(ggplot2)

ggplot(train, aes(x = Area, y = Perimeter)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Perimeter vs Area",
       x = "Area",
       y = "Perimeter")
```
Looks very linearly related. But quick linear regression to check:
```{r}
perim_area <- lm(Perimeter ~ Area, data = train)
summary(perim_area)
```
Very highly correlated.

## MajorAxisLength vs. MinorAxisLength scatterplot
```{r}
ggplot(train, aes(x = MinorAxisLength, y = MajorAxisLength)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "MajorAxisLength vs. MinorAxisLength",
       x = "MinorAxisLength",
       y = "MajorAxisLength")
```
### Quick analysis of MajorAxisLength vs. MinorAxisLength Linear Model
To assess strength of correlation
```{r}
major_minor <- lm(MajorAxisLength ~ MinorAxisLength, data = train)
summary(major_minor)
```
Seems relatively strong

## Area and MajorAxisLength, MinorAxisLength, ConvexArea, EquivDiameter
```{r}
ggplot(train, aes(x = MajorAxisLength, y = Area)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Area vs. MajorAxisLength",
       x = "MajorAxisLength",
       y = "Area")
ggplot(train, aes(x = MinorAxisLength, y = Area)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Area vs. MinorAxisLength",
       x = "MinorAxisLength",
       y = "Area")
ggplot(train, aes(x = ConvexArea, y = Area)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Area vs. ConvexArea",
       x = "ConvexArea",
       y = "Area")
ggplot(train, aes(x = EquivDiameter, y = Area)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Area vs. EquivDiameter",
       x = "EquivDiameter",
       y = "Area")
ggplot(train, aes(x = EquivDiameter, y = ConvexArea)) +
  geom_point(alpha = 0.6, col = "cadetblue") +
  theme_minimal() +
  labs(title = "Convexarea vs. EquivDiameter",
       x = "EquivDiameter",
       y = "Area")
```
```{r}
summary(lm(Area ~ EquivDiameter, data = train))
```
Area is super correlated with ConvexArea and EquivDiameter

*Solidity and roundness: NOT related

# Correlation Matrix

```{r}
correlation_matrix <- cor(train[, -which(names(train) == "Y")])
correlation_matrix
```
## Observations from correlation matrix:

1. `Area` and `Perimeter`: Correlation coefficient is 0.989 (strong positive correlation).
2. `Area` and `MajorAxisLength`, `MinorAxisLength`, `ConvexArea`, `EquivDiameter`: All have correlation coefficients above 0.94 with `Area`.
3. `Perimeter` and `MajorAxisLength`, `MinorAxisLength`, `ConvexArea`, `EquivDiameter`: Similar to `Area`, these variables also show strong positive correlations with `Perimeter`.

Based on observations 1, 2, and 3, we could potentially only look at one of the variables (e.g., Area) since they all seem to be very correlated.

4. `AspectRation` and `Eccentricity`: Correlation coefficient is 0.981.
5. `Compactness`, `ShapeFactor1`, `ShapeFactor3` and `AspectRation`, `Eccentricity`: These groups have correlation coefficients very close to -1 with each other, indicating strong negative correlations.

Conclusions based on 4 and 5?

```{r}
library(corrplot)
# Circles correlation plot
corrplot(cor(train), method = "circle",
        tl.pos = "n", mar = c(2, 1, 3, 1)) 
```

```{r, fig.width=19, fig.height=9}
# Visualization of correlation matrix
corrplot.mixed(cor(train),
               lower = "number", 
               upper = "circle",
               tl.col = "black")
```

# Outliers - Detection and Handling
```{r}
library(ggplot2)
ggplot(train, aes(x = ShapeFactor1)) + 
  geom_boxplot() +
    theme_minimal()
ggplot(train, aes(x = ShapeFactor2)) + 
  geom_boxplot() +
    theme_minimal()
ggplot(train, aes(x = ShapeFactor3)) + 
  geom_boxplot() +
    theme_minimal()
ggplot(train, aes(x = ShapeFactor4)) + 
  geom_boxplot() +
    theme_minimal()
```
*NOTE there are no missing data*

